Case Study: RNA-Seq Differential Analysis (basal ganglia)

### 2018-03-22 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Rename chipseq, rnaseq Rmd files

## intermediary files we create below
count_file <- file.path(resdir, "h3k4me3-csaw-counts.rds")
result_file <- file.path(resdir, "h3k4me3-csaw-results.rds")
bench_file <- file.path(resdir, "h3k4me3-csaw-benchmark.rds")

@@

```{r}
std_chr <- paste0("chr", c(1:19, "X", "Y"))
param <- readParam(minq = 15, discard = bl, restrict = std_chr)
```

@@

```{r count, results = 'hide', message = FALSE, warning = FALSE}
if (file.exists(count_file)) {
    filtered_cnts <- readRDS(count_file)
} else {
    ## Broad samples have quality score w/ phred offset +33 (standard)
    b_meta <- meta[meta$lab == "Broad", ]
    b_unaligned <- !file.exists(b_meta$bamfile)
    if (any(b_unaligned)) {
        align(index = file.path(datdir, "ref_index"),
              readfile1 = b_meta$fqfile[b_unaligned],
              TH1 = 2, type = 1, phredOffset = 33,
              output_file = b_meta$bamfile[b_unaligned])
    }

@@

    ## index bam files
    indexBam(meta$sortedfile)

@@

    ## filter windows by abundance
    bins <- windowCounts(meta$sortedfile, bin = TRUE, width = 2000, param = param)
    filter_stat <- filterWindows(win_cnts, bins, type = "global")
    keep <- filter_stat$filter > log2(3)
    filtered_cnts <- win_cnts[keep, ]

    saveRDS(filtered_cnts, file = count_file)
}
```

@@

We use _edgeR_ to test for differential binding.

@@

    saveRDS(res_ranges, file = result_file)
}
```

>>>>

## intermediary files we create below
count_file <- file.path(resdir, "h3k4me3-csaw-counts.rds")
filtered_file <- file.path(resdir, "h3k4me3-csaw-counts-filtered.rds")
result_file <- file.path(resdir, "h3k4me3-csaw-results.rds")
bench_file <- file.path(resdir, "h3k4me3-csaw-benchmark.rds")

@@

```{r}
std_chr <- paste0("chr", c(1:19, "X", "Y"))
param <- readParam(minq = 20, discard = bl, restrict = std_chr)
```

@@

```{r count, results = 'hide', message = FALSE, warning = FALSE}
if (file.exists(count_file)) {
    win_cnts <- readRDS(count_file)
} else {
    ## Broad samples have quality score w/ phred offset +33 (standard)
    b_meta <- meta[meta$lab == "Broad", ]
    b_unaligned <- !file.exists(b_meta$bamfile)
    if (any(b_unaligned)) {
        align(index = file.path(datdir, "ref_index"),
              readfile1 = b_meta$fqfile[b_unaligned],
              TH1 = 2, type = 1, phredOffset = 33,
              output_file = b_meta$bamfile[b_unaligned])
    }

@@

    ## index bam files
    if (any(!file.exists(paste0(meta$sortedfile, ".bai")))) {
        indexBam(meta$sortedfile)
    }

@@

    ## save unfiltered counts
    saveRDS(win_cnts, file = count_file)
}
```

We can apply prefiltering on windows with low abundance as described in Lun and Smyth (2016).

```{r}
if (file.exists(filtered_file)) {
    filtered_cnts <- readRDS(filtered_file)
} else {
    ## filter windows by abundance
    bins <- windowCounts(meta$sortedfile, bin = TRUE, width = 2000, param = param)
    filter_stat <- filterWindows(win_cnts, bins, type = "global")
    keep <- filter_stat$filter > log2(2)
    filtered_cnts <- win_cnts[keep, ]

    ## save filtered counts
    saveRDS(filtered_cnts, file = filtered_file)
}
```

@@

We use _edgeR_ to test for differential binding with the filtered data.

@@

    saveRDS(res_ranges, file = result_file)
}

## covert to df for downstream analysis
res_df <- as.data.frame(res_ranges)
```

## Covariate Diagnostics

### Number of Windows

```{r, nwin-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "nWindows") +
    ggtitle("Number of windows as independent covariate") +
    xlab("Number of Windows")
```

```{r, nwin-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 35, numQ = 5)
```

### Region Width

```{r, width-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "width") +
    ggtitle("Region width as independent covariate") +
    xlab("Region Width")
```

```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 5)
```

## Multiple-Testing Correction

We use the common `BenchDesign` with the set of multiple testing correction
methods already included. We use the width of the regions as the independent
covariate.

```{r}
if (!file.exists(bench_file)) {
    res_df$ind_covariate <- res_df$width
    res_df$pval <- res_df$PValue

    bd <- initializeBenchDesign()
    sb <- buildBench(bd, data = res_df, ftCols = "ind_covariate")
    saveRDS(sb, file = bench_file)
} else {
    sb <- readRDS(bench_file)
}
```

### 2018-03-22 datasets/RNA-Seq/RNAseq-gtex.Rmd
Rename chipseq, rnaseq Rmd files

We use script `downloadAndSubset.R` to pre-process the data. This script downloads a DESeqDataSet object containing gene level counts for a subset of the GTEx data. Then, it further subsets it keeping samples from only females and two cell types,  nucleus accumbens and the putamen.

```{r, dsd-download}
if (!file.exists(count_file)) {
    download.file("https://www.dropbox.com/s/pptxqnpsmdkyhcf/brain-counts.rds?dl=1", destfile=count_file)
}
dsd <- readRDS(count_file)
```

>>>>

We download a DESeqDataSet object containing gene level counts for a subset of the GTEx data. The data set includes  females and two cell types, nucleus accumbens and the putamen.

```{r, dsd-download}
if (!file.exists(count_file)) {
    download.file("https://www.dropbox.com/s/pptxqnpsmdkyhcf/brain-counts.rds?dl=1",
                  destfile = count_file)
}
dsd <- readRDS(count_file)
```
